RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-05-23 abstract.tex
Small modifications
Modified and commented on abstract and intro
Need to address the comments

\begin{abstract}
Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks. The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders. While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model. For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured. In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling. We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL). We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
\end{abstract}

>>>>

\begin{abstract}
Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks. The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders. While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found in the resulting model. For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured. In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to a goal modelingn approach. We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL). We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
\end{abstract}
%SG: what are the findings? or what case study we are using? 
%SG_New: I still have these comments!

### 2017-05-23 introduction.tex
background
checked the intro
Modified and commented on abstract and intro
Need to address the comments

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time. \todo{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
\item
Other stakeholders, such as new developers on the team, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders. %~\cite{castro2002towards}. %TODO Marc: if you like you can add this reference but I think it is not necessary. 
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
In goal models in general, ``rationale'' behind any decision is static and does not immediately impact the goal models when they change. That is, it is not possible to reason about changing beliefs and opinions, and their effect on the goal model. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying beliefs and arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
\end{enumerate}

The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with practical reasoning and argumentation approaches from Artificial Intelligence (AI)~\cite{atkinson2007}. We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must be able to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have tool support; (5) a methodology must be identified that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise %TODO Sepi_Reply: we need to come back to this at the end of the paper. 
 \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc. allows us to find rationales that would otherwise not have been found -- for this we would need an experiment. Tailor it towards our empirical work: ``the framework should be close to on the actual discussions stakeholders or designers have in the early requirements engineering phase''}.

@@

The contributions of this paper are aligned with our five success criteria. In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}. By combining arguments and critical questions, one can create a structured representation of a discussion. We, first, propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modeling process. We, then, validate this initial set with transcripts containing discussions about the architecture of an information system. This, then, gives us the final set of arguments and critical questions.

In order to generate goal models based on formalized discussions (requirement 2), we, first, formalize the list of arguments from requirement 1 in an argumentation framework. We formalize the critical questions as algorithms modifying the argumentation framework. We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected. We propose an algorithm to generate a GRL model based on accepted arguments. This creates traceability links from GRL elements to the underlying arguments (requirement 3).

We implement our framework in an online tool called RationalGRL (requirement 4). The tool is implemented using Javascript. It contains  two parts, goal modeling and argumentation. The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators. The argumentation part is new, and we develop a modeling language for the arguments and critical questions. The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the Eclipse-based tool for GRL modeling, for further evaluation and analysis.

@@

The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.

The rest of the article is in line with the main contributions and the research question. We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme. In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions. In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}. In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.

>>>>

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time. 
\todo[resolved]{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
% [Marc] We don't use preferences.
\item
Other stakeholders, such as new developers on the team, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders. %~\cite{castro2002towards}. %TODO Marc: if you like you can add this reference but I think it is not necessary. 
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
In goal models in general, ``rationale'' behind any decision is static and does not immediately impact the goal models when they change. That is, it is not possible to reason about changing beliefs and opinions, and their effect on the goal model. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying beliefs and arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
\end{enumerate}

The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with practical reasoning and argumentation approaches from Artificial Intelligence (AI)~\cite{atkinson2007}. We have identified several important requirements for our framework: (1) it must have the means to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must have the means to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have a tool support; (5) there should be a methodology that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise %TODO Sepi_Reply: we need to come back to this at the end of the paper. 
 \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc. allows us to find rationales that would otherwise not have been found -- for this we would need an experiment. Tailor it towards our empirical work: ``the framework should be close to on the actual discussions stakeholders or designers have in the early requirements engineering phase''}. %TODO SG: What about 6? we still don't have it!

@@

The contributions of this paper are aligned with our five %TODO SG: why 5?Not 6?
 success criteria. In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}. By combining arguments and critical questions, one can create a structured representation of a discussion. We, first, propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modeling process. We, then, validate this initial set with transcripts containing discussions about the architecture of an information system. This, then, gives us the final set of arguments and critical questions.

In order to generate goal models based on formalized discussions (requirement 2), we, first, formalize the list of arguments from requirement 1 in an argumentation framework. We formalize the critical questions as algorithms modifying the argumentation framework. We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected. We propose an algorithm to generate a GRL model based on the accepted arguments. This helps providing traceability links from GRL elements to the underlying arguments (requirement 3).

We implement our framework in an online tool called RationalGRL (requirement 4). The tool is implemented using Javascript. It contains  two parts, goal modeling and argumentation. The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators. The argumentation part is new, and we develop a modeling language for the arguments and critical questions. The created GRL models in RationalGRL can be exported to jUCMNav~\cite{} %TODO SG: add citation 
 the Eclipse-based tool for GRL modeling, for further evaluation and analysis.

@@

The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{}  %TODO SG: add citation
and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.

The rest of the article is in line with the main contributions and the research question. We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme. In Section~\ref{sect:gmas:2} %TODO SG: where is this?
, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions. In Section~\ref{sect} %TODO SG:which section?
, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}. In Section~\ref{sect:methodology} %TODO SG: where is this?
, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}. %TODO SG: where is this?

%TODO SG: In section 3, we talk about the methodology and then you discuss it again in section 7? why?
